---
title: "Example Report"
author: "Suzanne Dufault, PhD"
date: "2021-06-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)

# Packages I like to use
# install.packages("tidyverse") # Uncomment this line if you need to install the tidyverse package
library(tidyverse) # Good for general data wrangling. More info: https://www.tidyverse.org

# install.packages(c("knitr", "kableExtra")) 
library(knitr)
library(kableExtra) # For making prettier tables. More info: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
```

## R Markdown

This is an R Markdown file. The benefits of working out of a document like this is that you can easily write in normal text *and* run code. It's a good place to keep track of the code that you need and any annotation or thoughts on the results. It's like a Stata .do file, but prettier and a bit more powerful. 

## Data

In this first code "chunk", I am going to call in the data sample and look at the first 5 rows. 

```{r data}
# Call in the data
load("2021-06-30_awed-spatial-sample.RData")

# Look at the first 5 lines
head(awed_slice, n = 5) %>% 
  # Make a pretty table
  kable() %>% 
  kable_styling()
```

Here are some simple summaries:

```{r simple-summary}
# The number of test-negatives (dengue = 0) and test-positives (dengue = 1) in the dataset
table(awed_slice$dengue) 
```
There are 1,973 test-negatives and 130 test-positives in this sample of the dataset.

```{r simple-summary-2}
# This is kind of an ugly summary, but it gives you the basic information
summary(awed_slice)
```

In the summary, we see that in this sample of the AWED data, there are 11 variables. 

+ participant_id_shuffled - a character vector containing the shuffled participant ids
+ enrolment_date - the date of enrolment. First enrolled participant was on January 8, 2018. Last enrolled participant was enrolled on March 18, 2020.
+ illness_onset - the reported date of illness onset. First illness onset was January 4, 2018. Last illness onset was March 16, 2020.
+ cluster - the AWED study cluster containing the participant's home address (labelled 1-24)
+ intervention - whether the cluster was untreated (intervention = 0) or treated with wMel-carrying Ae. aegypti deployments (intervention = 1)
+ dengue - whether the participant tested negative for dengue (dengue = 0) or tested positive for dengue (dengue = 1)
+ longitude - of home location
+ latitude - of home location
+ serotype - either test-negative, the identified dengue serotype (DENV1-DENV4), or unknown dengue serotype (unable to identify)
+ age_shuffled
+ sex_shuffled

## Maps

Here, the code gets a bit complicated so I have hidden it from the output for now. Basically, I'm tidying up the data so that it will make a pretty map.

```{r map-prep, echo = FALSE, message=FALSE}
# install.packages("tmap") 
# install.packages("spatstat")
library(tmap)
library(rgdal)

# Setting up a color palette for the different DENV serotypes
df_col_sero <- data.frame(serotype = c(paste0("DENV", 1:4), "Unknown", "Test-negative"),
                          sero.color = c("#0AA4D1", "#F3C73E", "#EF7822", "#76C1A8", "#003D58", "gray"),
                          sero.shape = c(21:24, 4, 25))

# Tidying up data for mapping
work_ts_spat <- awed_slice %>% 
  mutate(serotype = case_when(serotype == "denv1" ~ "DENV1",
                              serotype == "denv2" ~ "DENV2",
                              serotype == "denv3" ~ "DENV3",
                              serotype == "denv4" ~ "DENV4",
                              serotype == "unk_serotype"~ "Unknown",
                              serotype == "test-negative control" ~ "Test-negative"),
         intervention = ifelse(intervention == 0, "Untreated", "Intervention")) %>% 
  full_join(df_col_sero) 

# Getting the "shape file" for the study area
Yogya_Adm_1 <- readOGR("yogya-shape-files/", "Yk_CaseControl_20160512")
int_assignment <- work_ts_spat %>% 
  dplyr::select(cluster, Intervention = intervention) %>% 
  distinct() %>% 
  mutate(New_Clustr = ifelse(cluster < 10, paste0("0", cluster), as.character(cluster))) %>% 
  dplyr::select(-cluster)
Yogya_Adm_1@data <- left_join(Yogya_Adm_1@data, int_assignment, by = "New_Clustr")

# Creating Spatial Points Dataframe 
# This is so that the points from our dataset can be mapped onto the study area map
CRTND_dengue_data_SPDF <- SpatialPointsDataFrame(coords = work_ts_spat[,c("longitude", "latitude")],
                                                 data = work_ts_spat[, c("cluster", "intervention", "enrolment_date", "illness_onset", "dengue", "age_shuffled", "sex_shuffled", "serotype", "sero.color", "sero.shape")],
                                                 proj4string = CRS("+init=epsg:4326")) # sets the projection to WGS 1984 using lat/long. Optional but good to specify

# Setting up a dataset with just the dengue cases
CRTND_VCD_data_SPDF <- subset(CRTND_dengue_data_SPDF, dengue == 1)
```

```{r map-prelim}
# A look at the preliminary map of the study area
map_yogya <- 
  tm_shape(Yogya_Adm_1) +
  tm_borders()

map_yogya + 
  tm_text(text = "New_Clustr")
```

```{r maps, fig.width=6, fig.height=7, fig.align='center'}
# Dengue point plot
map_yogya +
  tm_polygons(col = "Intervention",
              alpha = 0.5,
              palette = c("lightgray", "white")) +
  tm_shape(subset(CRTND_dengue_data_SPDF, dengue == 1)) + 
  tm_symbols(col = "serotype",
          shape = "serotype",
          legend.col.show = FALSE,
          legend.shape.show = FALSE,
          palette = df_col_sero$sero.color,
          shapes = df_col_sero$sero.shape,
          size = 0.15) + 
  tm_add_legend("symbol", 
                col = df_col_sero$sero.color[1:5], 
                shape = df_col_sero$sero.shape[1:5],
                labels = df_col_sero$serotype[1:5]) +
  tm_layout(legend.text.size = 0.9,
            legend.title.size = 1,
            legend.title.color = "white",
            inner.margins = 0.1,
            title = "Dengue Cases",
            title.fontface = "bold") +
  tm_scale_bar(breaks = c(0,.50,1.00), text.size = 0.9, position = c("left", "top"))

# Test-negative point plot
map_yogya +
  tm_polygons(col = "Intervention",
              alpha = 0.5,
              palette = c("lightgray", "white")) +
  tm_shape(subset(CRTND_dengue_data_SPDF, dengue == 0)) + 
  tm_symbols(shape = 3,
             col = "steelblue3",
          size = 0.15) + 
  tm_layout(legend.text.size = 0.9,
            legend.title.size = 1,
            legend.title.color = "white",
            inner.margins = 0.1,
            title = "Test-negatives",
            title.fontface = "bold") +
  tm_scale_bar(breaks = c(0,.50,1.00), text.size = 0.9, position = c("left", "top"))
```
